---
title: "Survival Analysis Assignment"
author: "Nishanth Girish"
date: "2023-09-16"
output: word_document
---


Consider the dataset Data.xlsx in which two methadone clinics for heroin addicts are
compared. The patient’s survival time (in days) is the amount of time the patient spent at the clinic before dropping out. In the dataset, the variables are defined as: 
ID - Patient ID,
SURVT - The time in days until the patient dropped out of the clinic or was censored
(missing information), STATUS - 1 for patient dropped out of the clinic or censored; 0
otherwise, 
CLINIC - Methadone Treatment Clinic Number 1 or 2,
PRISON - An indicator whether the patient had a prison record. 1 for yes, 0 for no, DOSE - Patient’s maximum methadone does (mg/day, continuous variable).

```{r}
# Loading the necessary libraries:

library(survival)
library(readxl)
```


```{r}
# Reading the data:

data = read_excel(file.choose())
attach(data)
data
```
```{r}
# Converting to a data frame:

data = as.data.frame(data)
```

```{r}
#a. Find the survival estimate for the survival variable and find the median time.

# Kaplan-Meier estimate:
result.km = survfit(Surv(survt, status) ~ 1, conf.type = "log-log" ) #1: no covariates
summary(result.km)

# Potential median survival time:
median(Surv(survt, status))     #Censoring is considered.

# Plot:
plot(result.km, conf.int = T, mark.time = T, xlab = "Time in days", ylab = "Survival probablity")
abline(h = 0.5, lty = 3, col = "Brown")
abline(v = 504, lty = 3, col = "Green") 
```

The potential median of the survival time is 504 days.

```{r}
#b. Check whether there is a difference in survival times among PRISON variable.

#[1: subject had a prison record, 0:subject does not have a prison record]
#H0: No significant difference in survival times among the 2 PRISON groups.
#H1: There is a significant difference in survival times among the 2 PRISON groups.

survdiff(Surv(survt, status) ~ prison)


#Plot:
plot(survfit(Surv(survt)~prison), xlab ="Time in days", 
     ylab ="Survival Probability", col =c("Blue", "Red"), lwd = 2)
legend("topright", legend = c("No Prison Record","Prison Record"), col = c("Blue","Red"),lwd=2)
```

#INTERPRETATION:

The p-value > 0.05. Thus, we fail to reject the null hypothesis. Thus, the 2 groups are not significantly different i.e there is no difference in survival times among the subjects who went to prison and subjects who did not go to prison. 

However, from the plot we can see visually that there is a difference among the 2 groups.

```{r}
#c. Fit a Cox proportional hazard model with CLINIC and PRISON as the covariates.

coxph(Surv(survt, status) ~ clinic + prison + dose)
```

#INTERPRETATION:

For CLINIC covariate, the p-value < 0.05. It is statistically significant. 
The beta is -1.1089 i.e unit increase in the CLINIC covariate results in -1.1089 unit increase in the log-hazard.

For PRISON covariate, the p-value > 0.05. It is not statistically significant. 
The beta is 0.2778 i.e unit increase in the PRISON covariate results in 0.2778 unit increase in the log-hazard.

The overall p-value of the likelihood ratio test is <0.05, indicating that the test is statistically significant.

[Cox doesn't have intercept as the baseline cancels out]

```{r}
survreg(Surv(survt, status) ~ clinic + prison + dose, data = data, dist = "weibull")
```

#INTERPRETATION:

p < 0.05, statistically significant. 

