---
title: "Untitled"
author: "Nishanth Girish"
date: "2023-08-23"
output: word_document
---

```{r}
library(survival)
```

```{r}
tt <- c(10,7,32,23,22,6,16,34,32,25,11,20,19,6,17,35,6,13,9,6,10)
cens <-c(1,1,0,  1, 1,1, 1, 0, 0, 0, 0, 0, 0,1, 0, 0,1, 1,0,0,0)

#to incorporate censored observations
Surv(tt, cens)
```


```{r}
result.km = survfit(Surv(tt, cens) ~ 1, conf.type = "log-log" ) #1: no covariates
summary(result.km)
```

```{r}
plot(result.km, xlab = "Time", ylab = "Survival probabilities", conf.int = T, mark.time = T, lty = 1:2, col = 1:3)
legend("bottomleft", legend = c("Estimates","CI"), cex = 1.0)#, lty = 1:2, col = 1:3)
```


```{r}
# For Nelson-Aalen estimator
result.km<-survfit(Surv(tt,cens)~1, conf.type = "log-log", type = "fh")
summary(result.km)

#lines(result.km, col = "red")
# From the plot we can say
```

##############################################################################################

```{r}
library(survival)
library(asaur)
head(gastricXelox)
data = gastricXelox
```
```{r}
data
```

```{r}
timeMonths<-data$timeWeeks*7/30.25
delta<-data$delta
result.km<-survfit(Surv(timeMonths, delta)~1, conf.type = "log-log")
summary(result.km)
          
```
```{r}
plot(result.km, conf.int = T, mark.time = T)
```
```{r}
x= seq(0,60,by=0.01)
hist(timeMonths , freq = F, ylim = c(0, 0.08))  #freq = F
lines(x, dweibull(x, shape = 1.09, scale =14.37),col ="red")
lines(x, dexp(x, rate = 0.072), col="green")
```

```{r}
library(fitdistrplus)
fitdist(timeMonths,"weibull" ,method= "mle")

fitdist(timeMonths, "exp", method= "mle")
```


```{r}
#survival = 1 - CDF
1- pweibull(timeMonths,shape=1.0998,scale=14.3710)

1 -pexp(timeMonths, rate = 0.0723)
```

```{r}
plot(1- pweibull(timeMonths,shape=1.0998,scale=14.3710))
```


```{r}
#which is the best fit? # KS test:
ks.test(timeMonths, pweibull,1.0998, 14.371)

ks.test(timeMonths, pexp, 0.0723)
```


```{r}
plot(result.km, conf.int = T, mark.time = T, xlab = "Time in months", ylab = "Survival probablity")
title("Progression-free survival in Gastric Cancer")
abline(h = 0.5, lty = 3, col = "Brown")
abline(v = 10.2975, lty = 3, col = "Green") 
```
```{r}
median(timeMonths)
# Simple median follow-up time, censoring not considered
median(Surv(timeMonths, delta))
# Potential median follow up time

result.km

plot(result.km, conf.int = T, mark.time = T, cumhaz = T, xlab = "Time in Months", ylab = "Cumulative Hazard")
```

##############################################################################################



```{r}
library(asaur)
head(pancreatic)
```
##############################################################################################


#Left Truncation
```{r}
library(survival)
library(asaur)
head(ChanningHouse)
tail(ChanningHouse)
```
```{r}
data = ChanningHouse
data
```

```{r}
ChanningHouse <- within(ChanningHouse, {
    entryYears <- entry / 12
    exitYears <- exit/12})
ChanningMales <- ChanningHouse[ChanningHouse$sex == "Male",]


result.km <- survfit(Surv(entryYears, exitYears, cens, type= "counting")~1 , data = ChanningMales, conf.type ="none")
summary(result.km)
plot(result.km, xlim= c(64,101), xlab= "Age", ylim=c(0,1), ylab = "Survival Probability", conf.int = F)


#Nelson-Aalen Estimator
result.nae <- survfit(Surv(entryYears, exitYears, cens, type= "counting")~1 , type = "fleming-harrington", data = ChanningMales)
summary(result.nae)
lines(result.nae, col="blue", conf.int=F)

result.km.68 <- survfit(Surv(entryYears, exitYears, cens, type= "counting")~1 , start.time = 68, data = ChanningMales)
summary(result.km.68)
lines(result.km.68, col ="green", conf.int = F)

```
```{r}
library(survival)
library(asaur)
```


```{r}
#Log-rank test:

tt <- c(6, 7, 10, 15, 19, 25)
delta <- c(1, 0, 1, 1, 0, 1)
trt <- c(0, 0, 1, 0, 1, 1)
survdiff(Surv(tt, delta) ~ trt)
```
#Interpretation:
p>0.05 (Calculated value < table value). We fail to reject the null hypothesis. Thus, the 2 groups are not significant different. There is no effect of the treatment.

##############################################################################################

```{r}
library(survival)
library(asaur)
```


```{r}
data <- pancreatic
data
```
```{r}
data$onstudy=as.Date(data$onstudy,format="%m/%d/%Y")
data$onstudy
data$death=as.Date(data$death,format="%m/%d/%Y")
data$death
data$progression=as.Date(data$progression,format="%m/%d/%Y")
data$progression

```
```{r}
data
```
```{r}
pancreatic2
```

```{r}
pfs = data$progression - data$onstudy
os = data$death - data$onstudy

pfs
os
```
```{r}
data = pancreatic2
attach(data)
survdiff(Surv(pfs, status ) ~ stage, rho  = 1)
```
#Interpretation:
p>0.05 (Calculated value < table value). We fail to reject the null hypothesis. Thus, the 2 groups are not significant different. 

```{r}
pfs.month = pfs/30.5
plot(survfit(Surv(pfs.month)~stage), xlab ="Time months", 
     ylab ="Survival Probability", col =c("Blue", "Red"), lwd = 2)
legend("topright", legend = c("Locally Advanced(LA)","Metastatic(M)"), col = c("Blue","Red"),lwd=2)
```
##############################################################################################
```{r}
library(survival)
library(asaur)
```


```{r}
data = pharmacoSmoking
attach(data)
data
```


```{r}
survdiff(Surv(ttr, relapse ) ~ grp)
```
#Interpretation:
p < 0.05 (Calculated value > table value). We reject the null hypothesis. Thus, the 2 groups are significant different. 

```{r}
plot(survfit(Surv(ttr)~grp), xlab ="Time in days", 
     ylab ="Survival Probability", col =c("Blue", "Red"), lwd = 2)
legend("topright", legend = c("Combination","Patch only"), col = c("Blue","Red"),lwd=2)
```
```{r}
#STRATIFIED LOG-RANK TEST:
survdiff(Surv(ttr, relapse ) ~ grp + strata(ageGroup2))
```
#Interpretation:
p < 0.05 (Calculated value > table value). We reject the null hypothesis. Thus, the 2 groups are significant different. 
```{r}
plot(survfit(Surv(ttr)~grp + strata(ageGroup2)), xlab ="Time in days", 
     ylab ="Survival Probability", col =c("Blue", "Red"), lwd = 1)
legend("topright", legend = c("Combination","Patch only"), col = c("Blue","Red"),lwd=1)
```

```{r}
library(survival)
library(asaur)
```
```{r}
tt <- c(6, 7, 10, 15, 19, 25)
delta <- c(1, 0, 1, 1, 0, 1)
trt <- c(0, 0, 1, 0, 1, 1)
```
```{r}
coxph(Surv(tt, delta) ~ trt)
```

p value > 0.05. Fail to reject null hypothesis. 
hazard rate is lower for experimental group when compared to the control group. Not significant. 